---
title: "Team R-Mageddon"
author: 
  - Ethan Allavarpu
  - Dara Tan
subtitle: 
date: |
  | ethanallavarpu@g.ucla.edu , tanxuendara@ucla.edu
  | Spring 2021
output:
  beamer_presentation:
    theme: "Hannover"
    colortheme: "spruce"
    includes:
      in_header: header.tex
    slide_level: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
load("helpgraph.RData")
source("state_politics.R")
source("two_var_relationship.R")
library(tidyverse)
library(randomForest)
```

# Drugs with Misuse Problems

## Income-Independent

```{r, echo = FALSE, cache = TRUE}
income
```

## Across the United States

```{r, message = FALSE, echo = FALSE, cache = TRUE}
source("state_nmu.R")
source("state_nmu_prop.R")
state_nmu[is.na(state_nmu)] <- 0
crazy <- c("COD", "HYD", "OXY", "BENZ", "STIM") %>% tolower()
for (i in seq_len(ncol(states) - 1) + 1) {
  if(tolower(names(states)[i]) == crazy[1] ) {print(plot_usmap(data = states, 
             values = names(states)[i],  # Change this for respective variable 
             color = rgb(0, 0, 0), 
             labels = FALSE) + 
    scale_fill_continuous( low = "white", 
                           high = rgb(0.5, 0, 0), 
                           name = "Average NMU", 
                           label = scales::comma,
                           limits = if(i < 19) {range(states[, -c(1, 19, 20, 21)])}else{NULL}) + 
    theme(legend.position = "right") + 
    theme(panel.background = element_rect(color = "black")) + 
    labs(title = paste("Average", toupper(names(states)[i]), "NMU by State", collapse = " ")))}
}
```

## Across the United States

```{r, message = FALSE, echo = FALSE, cache = TRUE}
for (i in seq_len(ncol(states) - 1) + 1) {
  if(tolower(names(states)[i]) == crazy[2] ) {print(plot_usmap(data = states, 
             values = names(states)[i],  # Change this for respective variable 
             color = rgb(0, 0, 0), 
             labels = FALSE) + 
    scale_fill_continuous( low = "white", 
                           high = rgb(0.5, 0, 0), 
                           name = "Average NMU", 
                           label = scales::comma,
                           limits = if(i < 19) {range(states[, -c(1, 19, 20, 21)])}else{NULL}) + 
    theme(legend.position = "right") + 
    theme(panel.background = element_rect(color = "black")) + 
    labs(title = paste("Average", toupper(names(states)[i]), "NMU by State", collapse = " ")))}
}
```

## Across the United States

```{r, message = FALSE, echo = FALSE, cache = TRUE}
for (i in seq_len(ncol(states) - 1) + 1) {
  if(tolower(names(states)[i]) == crazy[3] ) {print(plot_usmap(data = states, 
             values = names(states)[i],  # Change this for respective variable 
             color = rgb(0, 0, 0), 
             labels = FALSE) + 
    scale_fill_continuous( low = "white", 
                           high = rgb(0.5, 0, 0), 
                           name = "Average NMU", 
                           label = scales::comma,
                           limits = if(i < 19) {range(states[, -c(1, 19, 20, 21)])}else{NULL}) + 
    theme(legend.position = "right") + 
    theme(panel.background = element_rect(color = "black")) + 
    labs(title = paste("Average", toupper(names(states)[i]), "NMU by State", collapse = " ")))}
}
```

## Across the United States

```{r, message = FALSE, echo = FALSE, cache = TRUE}
for (i in seq_len(ncol(states) - 1) + 1) {
  if(tolower(names(states)[i]) == crazy[4] ) {print(plot_usmap(data = states, 
             values = names(states)[i],  # Change this for respective variable 
             color = rgb(0, 0, 0), 
             labels = FALSE) + 
    scale_fill_continuous( low = "white", 
                           high = rgb(0.5, 0, 0), 
                           name = "Average NMU", 
                           label = scales::comma,
                           limits = if(i < 19) {range(states[, -c(1, 19, 20, 21)])}else{NULL}) + 
    theme(legend.position = "right") + 
    theme(panel.background = element_rect(color = "black")) + 
    labs(title = paste("Average", toupper(names(states)[i]), "NMU by State", collapse = " ")))}
}
```


## Across the United States

```{r, message = FALSE, echo = FALSE, cache = TRUE}
for (i in seq_len(ncol(states) - 1) + 1) {
  if(tolower(names(states)[i]) == crazy[5] ) {print(plot_usmap(data = states, 
             values = names(states)[i],  # Change this for respective variable 
             color = rgb(0, 0, 0), 
             labels = FALSE) + 
    scale_fill_continuous( low = "white", 
                           high = rgb(0.5, 0, 0), 
                           name = "Average NMU", 
                           label = scales::comma,
                           limits = if(i < 19) {range(states[, -c(1, 19, 20, 21)])}else{NULL}) + 
    theme(legend.position = "right") + 
    theme(panel.background = element_rect(color = "black")) + 
    labs(title = paste("Average", toupper(names(states)[i]), "NMU by State", collapse = " ")))}
}
```



# Two-Variable Interactions

## Age and Race

```{r, message=FALSE, echo=FALSE}
two_var_relationship("DEM_AGE10", "DEM_RACE")
```

## Alcohol and Recreational Cannibis Use

```{r, message=FALSE, echo=FALSE}
two_var_relationship("ALC_USE", "CAN_REC_USE")
```

## Income and Education 

```{r, message=FALSE, echo=FALSE}
two_var_relationship("DEM_EDU", "DEM_DOLLARS")
```

## Income and Education

```{r, echo=FALSE, message=FALSE, cache = TRUE}
x <- "DEM_EDU"
y <- "DEM_DOLLARS"
  race_age <- list()
  length(race_age) <- 1
  names(race_age) <- "ANY"
  us18$var1 <- us18[[x]]
  us18$var2 <- us18[[y]]
  ra <- us18 %>%
    mutate(NMU = nmu_aggregate, ANY = any_nmu) %>%
    mutate_all(~replace(., is.na(.), 0)) %>%
    group_by(var1, var2) %>%
    summarize(dast = weighted.mean(ANY, WT, na.rm = TRUE))
  for (i in seq_len(ncol(ra) - 2)) {
    ra[is.na(ra)] <- 0
    ra_new <- as.data.frame(ra[, c(1:2, i + 2)])
    colnames(ra_new) <- c(x, y, "NMU")
    race_age[[i]] <- ra_new
  }
  par(mfrow = c(1, 3))
  for (i in seq_along(race_age)) {
    cor_heatmap <- ggplot(data = race_age[[i]], aes(x = race_age[[i]][, 1], y = race_age[[i]][, 2], fill = race_age[[i]][, 3])) +
      geom_tile()
    cor_heatmap <- cor_heatmap +
      scale_fill_gradient2(low = "white", high = "darkred",
                           name = "ANY") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1)) +
      ggtitle("Any Drug Misuse") + xlab(x) + ylab(y)
    print(cor_heatmap)
  }
```

## Chronic Pain and Education

```{r, message=FALSE, echo=FALSE}
two_var_relationship("PAIN_CHRONIC", "DEM_EDU")
```

## Prescriptions for Chronic and Acute Pain

```{r, message=FALSE, echo=FALSE}
two_var_relationship("PAIN_CHRONIC_RX", "PAIN_ACUTE_RX")
```

# Influential Variables

## Demographics
```{r, message=FALSE, echo=FALSE, cache = TRUE}
load("pval.RData")
temp_plot
```


## Politics

```{r, message=FALSE, echo=FALSE, cache = TRUE}
for (i in 19) {
    # political affiliation by 2018 election results ----
  print(plot_usmap(data = nmu_politics, values = "election",
                   labels = FALSE) +
          scale_fill_gradient2(low = rgb(0.5, 0, 0), mid = rgb(0.5, 0, 0.5, alpha = 0.1),
                               high = rgb(0, 0, 0.5),
                               name = "Political Affiliation") +
          theme(legend.position = "right") +
          theme(panel.background = element_rect(color = "black")) +
          labs(title = "Political Affiliation by 2018 Election Results"))
}
```

## Politics
```{r, message=FALSE, echo=FALSE, cache = TRUE}
    # average NMU by state ----
for(i in 19) {
  print(plot_usmap(data = nmu_politics,
                   values = names(nmu_politics)[i], labels = FALSE) +
          scale_fill_continuous(
            low = "white", high = "darkred", name = "Average NMU",
            limits = if (i < 19) {range(nmu_politics[, -c(1, 19:23)])
              } else {NULL}) +
          theme(legend.position = "right") +
          theme(panel.background = element_rect(color = "black")) +
          labs(title =
                 str_c("Average ", names(nmu_politics)[i], " NMU By State")))
}
```

## Politics

```{r, echo=FALSE, message=FALSE, cache = TRUE}
for (i in 2:19) {
  df <- nmu_politics %>% select(state, election, nmu = names(nmu_politics)[i])
  if (names(nmu_politics)[i] == "dast") {
  print(df %>% ggplot() +
          geom_col(aes(reorder(state, -nmu), nmu, fill = election)) +
          scale_fill_gradient2(low = rgb(0.5, 0, 0), mid = rgb(0.5, 0, 0.5, alpha = 0.1),
                               high = rgb(0, 0, 0.5),
                               name = "Political Affiliation") +
          theme_bw() +
          labs(title = str_c("Weighted Mean of ", toupper(names(nmu_politics)[i]),
                             " NMU by State and Political Affiliation"),
               x = "State", y = str_c("Weighted Mean of ",
                                      names(nmu_politics)[i], " NMU")))
  }
}
```




## Random Forest

```{r, echo = FALSE, cache = TRUE}
library(randomForest)
varImpPlot(bag_tree, type = 1, n.var = 10, main = "Influential Predictors for DAST Scores")
```


# Conclusion

## Misused Drugs

  - Hydrocodone (e.g., Vicodin)
  - Codeine
  - Oxycodone (e.g., OxyContin, Percocet)
  - Benzodiazepines (e.g., Xanax)
  - Stimulants (e.g. Adderall)
  

## How to Identify Misuse

  - Prior misuse is a strong indicator of future misuse (expected)
  - Age brackets seem to be important (younger people have higher levels of misuse)
  - Higher education levels surprisingly had higher levels of misuse for *any* drug
  - Acute pain prescriptions was a stronger driver of misuse than chronic pain prescriptions
  - Race is not as strong of a factor as many perceive

# Citations

  - https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/Information-on-Prescription-Drugs/MedicarePartD
  - https://github.com/MEDSL/2018-elections-official